\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.6cm]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{tabularx}
\usepackage{pdflscape}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta}
\onehalfspacing

\title{From Supply and Use Tables to Reproducible Social Accounting
Matrices:\\ A Transparent Construction and Audit Framework\\
\large Evidence from three publication regimes: South Africa, Kenya,
and the Netherlands --- working paper, draft v0.8}
\author{Kodzovi Abalo\thanks{World Bank. The views expressed here are those
of the author and do not necessarily reflect the views of the World Bank,
its Executive Directors, or the governments they represent. Comments
welcome.}}
\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
\noindent
Social accounting matrices (SAMs) underpin much applied policy modelling,
yet the way they are built has a reproducibility problem: heterogeneous
official tables are reconciled through undocumented spreadsheet operations
and expert judgments that cannot be audited, updated, or transferred. This
paper demonstrates that the problem is unnecessary---that SAM construction
can move from an artisanal exercise to an auditable statistical-production
process. We introduce a construction and audit framework in which every
source is fingerprinted, every classification and formula is
machine-readable, every balancing adjustment is logged at the cell where
it acts, and accounting identities are enforced as executable checks. We
apply it to build a complete 2019 SAM for South Africa from publicly
obtainable official sources and access-controlled survey microdata,
audited cell by cell against the published UNU-WIDER benchmark of
\citet{vanseventer2023}: the benchmark's production and commodity
structure is reproduced exactly, and its institutional macro structure is
reproduced exactly except for one unpublished input---separately
identified and priced ($+16.1$ per cent on import duties under its public
proxy)---and one bounded unrecorded reconciling adjustment. Where the benchmark's choices are
undocumented, the framework recovers them numerically; every reported
discrepancy was first detected by deterministic reconciliation and then
interpreted against the source record. Two further, deliberately
shallower applications test the framework at the ends of the public-data
spectrum: in Kenya, auditing proceeds exactly as far as the boundary set
by publication format; in the Netherlands, the framework generates and
balances a macro SAM from three Eurostat API calls with no benchmark
consulted. The open toolkit that implements the framework is offered as
reproducible infrastructure, and the central proposal follows: cell-level
provenance and machine-readable construction rules, published alongside
every SAM, would make such audits---and such SAMs---routine.
\end{abstract}

\vspace{0.5em}
\noindent\textbf{Keywords:} social accounting matrix, supply and use tables,
reproducibility, data provenance, statistical auditing, South Africa,
Kenya, Netherlands.\\
\textbf{JEL:} C67, C82, D57, E16, O55.

\input{sections/introduction}
\input{sections/related}
\input{sections/data}
\input{sections/method}
\input{sections/results}
\input{sections/applications}
\input{sections/discussion}
\input{sections/conclusion}
\input{sections/disclosure}

\bibliographystyle{apalike}
\bibliography{references}

\appendix
\input{sections/appendix}

\end{document}
