\section{Disclosed construction rules}\label{app:rules}

Table~\ref{tab:rules} reproduces the numbered rules under which our SAM
departs from the benchmark's (unpublished or unrecorded) choices. Each rule
is implemented in the replication package and printed in the generated
workbook; ``measured consequence'' refers to the comparison against the
benchmark reported in Section~\ref{sec:results}.

\begin{table}[h]
\centering\small
\caption{Rules R1--R7: explicit substitutes for non-public or unrecorded choices.}
\label{tab:rules}
\begin{tabularx}{\textwidth}{lXX}
\toprule
Rule & Construction & Replaces / measured consequence \\
\midrule
R1 & Import duties by commodity: total from the public series KBP4590J, allocated by import shares & Unpublished customs data; $+16.1$\% on import duties \\
R2 & Other product taxes by commodity: supply-table tax column minus R1 & Same dependency; $-1.5$\% on product taxes \\
R3 & Capital-type shares for six activities absent from the AFS: economy-wide average shares & Undocumented benchmark source for those sectors \\
R4 & Household consumption pattern: each commodity split across groups by total-expenditure shares & Benchmark's inherited item-level shares; drives most residual household imbalance \\
R5 & Occupation-to-household wages: aggregate LCS wage-income group shares for every occupation & Benchmark's education-based expansion \\
R6 & Other household-facing cells: LCS wage-income group shares as income-side proxy & Benchmark's item-specific LCS systems \\
R7 & Capital income to institutions uniform across the six capital types & The benchmark's own documented rule (adopted) \\
\bottomrule
\end{tabularx}
\end{table}

\section{Evidentiary status of the audit findings}\label{app:evidence}

Not every divergence between record and artefact has the same epistemic
status, and the audit methodology requires saying which is which.
Table~\ref{tab:evidence} classifies each South African finding by its
detection mechanism and the strength of the resulting inference, on a
five-level scale: \emph{exactly recovered} (the data admit one answer and
it reproduces the benchmark); \emph{unique within tested grid} (one
candidate among an enumerated finite set fits, with the runner-up margin
reported); \emph{consistent with} (the pattern matches a hypothesised
intervention that is not uniquely identified); \emph{documented} (the
source record itself states the choice); and \emph{unresolved} (the
inputs needed to decide are unavailable). The table has ten rows in total:
seven covering the eight record--artefact divergences catalogued in
Section~\ref{sec:discussion} (the two customs-related divergences---the
double-printed value and the non-public input---combined in one row),
plus three supplementary items reported for completeness: a recovered
grouping choice, an unresolved data gap, and a residual outlier.

\begin{landscape}
\begin{table}[p]
\centering\small
\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textheight}{%
  >{\raggedright\arraybackslash}p{4.6cm} X
  >{\raggedright\arraybackslash}p{4.8cm}}
\toprule
Finding & Detection mechanism & Evidentiary status \\
\midrule
\texttt{atx}, \texttt{cler}, \texttt{craf} misnamed & identity match on
all 61 activities & exactly recovered \\
\texttt{mach}/\texttt{nitc}/\texttt{trnp} rotation & dictionary
cross-check against technical note & documented \\
Formula signs (intra-enterprise cell) & exhaustive $3^6$ grid; runner-up
misses by R2.5bn & unique within tested grid \\
Capital-split vintage and measure & edition $\times$ measure grid
(0.18pp vs 0.72pp) & unique within tested grid \\
Household grouping variable & wage-income anchor (the consumption
anchor is uninformative: inherited shares) & unique within tested
definitions \\
R596m government--enterprise edit & balance-preserving minimal-edit
pattern; vintage test refuted & consistent with \\
Customs input (\texttt{mtx}/\texttt{stx}) & note's own statement;
priced by public proxy & documented; unresolved input \\
Six sectors' capital shares & absent from AFS source & unresolved \\
Household consumption shares & technical-note statement & documented
(inherited system) \\
Posts/telecoms capital outlier & residual after recovered basis &
consistent with (unreported reclassification) \\
\bottomrule
\end{tabularx}
\caption{South African audit findings by detection mechanism and
evidentiary status.}
\label{tab:evidence}
\end{table}
\end{landscape}

\section{Balancing the generated Dutch matrix: rules, invariants, and
sensitivity}\label{app:nlbalance}

The balancing step of Section~\ref{sec:applications:nl} involves three
choices, none mechanically inevitable; this note states each rule, the
invariant it preserves, and the measured sensitivity to the one genuinely
discretionary choice.

\emph{Sign normalisation.} The generated matrix contains two negative
cells: net property-type capital income vis-\`a-vis the rest of the world
(EUR $-10{,}286$m at (row $\leftarrow$ capital)) and the D8 pension
adjustment vis-\`a-vis the rest of the world (EUR $-168$m). Each is
flipped to its positive transpose before balancing: a payment of $-v$
from account $a$ to account $b$ is identically a payment of $+v$ from
$b$ to $a$. The flip preserves, cell by cell, every account's
receipts-minus-payments residual and every institution's net position;
it changes only the gross flows' side of the ledger, and both flips have
the natural economic reading (the Netherlands is a net \emph{recipient}
of capital income from abroad). RAS requires a non-negative seed, so the
alternative would be GRAS \citep{junius2003}; normalisation is preferred
\emph{at this scale} because it makes the negativity explicit and
reviewable, two cells being small enough to inspect individually. The
preference does not generalise: a full SAM carries structural negatives
throughout (subsidies, negative saving, inventory changes), and
transposing each one would rearrange the matrix away from the
conventional layout. For that case GRAS is the appropriate instrument,
and the toolkit implements it.

\emph{Target vector.} RAS needs one target per account. The generated
matrix supplies two candidates---each account's receipts and its
payments, which differ by exactly the attributed residual---and the rule
takes their arithmetic mean: the symmetric choice that favours neither
side of any account and whose row- and column-target totals agree by
construction. The two one-sided choices (all-receipts, all-payments) are
the extreme brackets of this decision.

\emph{Sensitivity.} Rebalancing against each extreme bracket (scaled to
a common total so variants are comparable cell by cell) moves no cell of
the balanced matrix by more than EUR~6.4~billion---0.9\,per\,cent of
GDP, the same order as the pre-balancing residuals themselves---and no
cell above EUR~100m by more than 6.6\,per\,cent of its value. That is
the expected behaviour: the target choice redistributes the attributed
residuals between the two sides of each unbalanced account; it does not
amplify them, and the mean-target SAM lies between the brackets.
Sensitivity is regenerated by the replication package alongside the
matrix.

\emph{Choice of RAS.} After normalisation the seed is non-negative, so
GRAS is unnecessary rather than rejected. RAS is preferred to
cross-entropy estimation here because its per-cell multiplicative
factors are individually interpretable and are published with the
matrix---the framework's logging commitment---while cross-entropy with
uninformative priors would deliver essentially the same adjustment with
less transparent diagnostics \citep{robinson2001}.

\section{Transfer validation across additional ESA publication regimes}
\label{app:transfer}

After the Dutch case was complete, the packaged generator
(\texttt{toolkit/recipes/build/} in the replication package) was rerun,
unchanged, on two further countries chosen to bracket the harmonised
system's publication practice; validation reports for both ship with the
replication package.

\emph{Austria 2019} behaves like the Netherlands: 65 industries and
products, zero identity findings at the EUR~1m tolerance, full commodity
closure, and every pre-balancing account residual below 0.6\,per\,cent
of GDP.

\emph{Poland 2019} exposed a publication-practice boundary \emph{inside}
the harmonised system: Polish tables withhold confidential cells, so the
published industry-and-product detail covers only 96.7\,per\,cent of
gross output (government final consumption detail: 92.6\,per\,cent). The
toolkit detects the suppression through its tiling and coverage
assertions, anchors the macro accounts to the published totals---which
remain complete and internally consistent---and reports the coverage of
every affected aggregate; the generated Polish matrix closes with
residuals below 1.1\,per\,cent of GDP. Even within the ESA transmission
programme, then, the replicability spectrum persists at the cell level:
harmonisation fixes the format, not the disclosure.

\section{The toolkit and replication package: a reader's guide}
\label{app:toolkit}

The toolkit (\texttt{edikit}, Apache-2.0; repository
\url{https://github.com/kodzovee9/tekmeris}, archived at
\href{https://doi.org/10.5281/zenodo.21419195}{doi:10.5281/zenodo.21419195})
is a small Python library in four modules---\texttt{data} (readers for supply--use workbooks, SAM
matrices, central-bank series, Eurostat JSON-stat, and concordances, each
enforcing its source's accounting identities on ingestion),
\texttt{model} (the account taxonomy with coverage assertions; RAS
balancing with mandatory per-cell factor diagnostics), \texttt{report}
(generation of workbooks whose totals and balance checks are live
formulas), and \texttt{pipeline} (end-to-end generation and audit
routines)---plus a test suite in which the accounting identities are the
test cases. Users enter through three documented \emph{recipes},
matching the publication regimes of this paper: \texttt{replicate/}
(regenerate every number here), \texttt{audit/} (audit a published SAM:
structure, then macro controls against published national accounts), and
\texttt{build/} (generate and balance a macro SAM for any
ESA-transmitting country). The latter two are each one command, e.g.:

{\small
\begin{verbatim}
python build_sam.py --country AT --year 2019
python audit_sam.py --sam <matrix.csv> --kinds <kinds.csv> \
                    --controls <controls.csv>
\end{verbatim}}

\noindent Step-by-step instructions live in the package's README files,
which can evolve with the code; this appendix instead gives the map a
verifying reader needs: which script produces each of the paper's
claims, and which validation report records it
(Table~\ref{tab:claimmap}). Every script writes its report to
\texttt{outputs/}; the reports, not the console, are the record.

\begin{table}[h]
\centering\small
\begin{tabularx}{\textwidth}{Xll}
\toprule
Claim & Script(s) & Validation report \\
\midrule
SUT identities pass; production and commodity blocks exact
  & 01--03 & \texttt{supply\_/use\_validation} \\
Institutional macro cells from SARB formulas; grid-recovered signs;
  priced customs proxy & 04--06 & \texttt{macro\_sam\_construction} \\
Capital split: vintage/measure recovery grid & 07 &
  \texttt{capital\_split} \\
Occupation shares from LMD wage bills & 08 & \texttt{occupation\_split} \\
Household grouping identification & 09 & \texttt{household\_split} \\
Full assembly under rules R1--R7; comparable-cell counts & 10 &
  \texttt{assembly\_report} \\
RAS balancing, factors logged; workbook & 11--12 &
  \texttt{balance\_report}; workbook \\
Paper exhibits regenerate from outputs & 13 & \texttt{paper/tables},
  \texttt{figures} \\
Kenya: structure, rounding grain, KNBS macro controls & 14 &
  \texttt{kenya\_structural\_audit} \\
Netherlands: generation, residual attribution, balancing and its
  sensitivity (Appendix~\ref{app:nlbalance}) & 15 &
  \texttt{netherlands\_generation} \\
Austria/Poland transfer validation (Appendix~\ref{app:transfer}) &
  build recipe & \texttt{\{AT,PL\}\_2019\_generation} \\
\bottomrule
\end{tabularx}
\caption{Map from the paper's claims to the scripts that produce them
and the validation reports that record them. Scripts 01--12 need only
the registered sources; 08--09 additionally require the DataFirst
microdata (Section~\ref{sec:data}).}
\label{tab:claimmap}
\end{table}
