\section{Data}\label{sec:data}

Every input is recorded in a source registry with publisher, vintage, access
date, licence status, and SHA-256 hash; raw files are never modified, and
artefacts derived from them carry lineage to the registry entry. The South
African sources are:

\begin{itemize}
  \item \textbf{Supply and use tables} (Stats~SA Report 04-04-03): the
  2018--2019 workbook with supply and use tables at 124 industries and 108
  products, plus the accompanying methodological report. These are the sole
  inputs to the production and commodity blocks.
  \item \textbf{Benchmark SAM} (UNU-WIDER Technical Note 2023/1): the
  distributed workbook of \citet{vanseventer2023}---two 209-account SAMs
  (occupational and educational wage detail), an account dictionary, and
  industry/product concordances---plus the technical note itself, from which
  we transcribe the 46 macro cell formulas and three mapping tables. The
  workbook is used \emph{only} as the audit target and as the published
  source of its own concordances; no benchmark value enters our SAM except
  where explicitly flagged (the single unpublished input, in the as-built
  scenario only).
  \item \textbf{SARB Quarterly Bulletin} (December 2022 bulk files): all 69
  annual KBP series required by the technical note's formulas. We verify the
  vintage directly: household consumption and imports in these files equal
  the benchmark's corresponding cells exactly, at the precision the
  bulletin files report (this vintage check is the one comparison made at
  source precision rather than at the R0.5~million audit tolerance of
  Section~\ref{sec:method:terms}).
  \item \textbf{Annual Financial Statistics} (Stats~SA P0021): the PPE
  schedules (asset type $\times$ SIC industry) from both the 2019
  preliminary and 2019 revised editions---the comparison matters, as
  Section~\ref{sec:results} shows.
  \item \textbf{Labour Market Dynamics} 2018 and 2019 and \textbf{Living
  Conditions Survey} 2014/15 microdata (Stats~SA via DataFirst): used only
  to compute aggregated shares (occupational wage bills; household groups).
  Under the access conditions, no microdata are redistributed; the
  replication package ships the scripts and the aggregated shares.
\end{itemize}

One input to the benchmark is not public: the split of import duties from
other product taxes rests on customs data that its technical note describes
as ``made available by Statistics South Africa as unpublished
data''.\footnote{\citet[item vii]{vanseventer2023}. The note does not
describe how the data were provided or in what form; we assert nothing
beyond its own wording.} We treat this input as a first-class object:
registered in the source registry, flagged in the formula table, propagated
to the three cells it touches, and priced by a public proxy
(Section~\ref{sec:results:macro}).

The two further applications add three registered sources. For Kenya: the
IFPRI 2019 Nexus SAM \citep{thurlow2021}, distributed under CC-BY-4.0 via
Harvard Dataverse, and the KNBS Economic Survey 2020
\citep{knbs2020}, whose current-price national-accounts tables supply the
macro controls; the 2009 benchmark supply and use tables exist in
condensed form in the Economic Survey 2014, with the full 151-product
database available from KNBS on request but not published
machine-readably. For the Netherlands: the Eurostat dissemination API
\citep{eurostat2024}, from which three requests retrieve the 2019 supply
table, use table, and non-financial sector accounts (datasets
\texttt{naio\_10\_cp15}, \texttt{naio\_10\_cp16},
\texttt{nasa\_10\_nf\_tr}); the retrieved JSON responses are registered
and hashed like every other raw input.
