\section{Discussion}\label{sec:discussion}

\subsection{A taxonomy of divergence between record and artefact}

Set side by side, the audit's findings sort into five classes, and it is
worth stating plainly that every reported finding was \emph{first
detected} by arithmetic---an identity refusing to close, a grid search
returning a unique survivor---and only then interpreted against the
source record (Appendix~\ref{app:evidence} states each finding's
evidentiary status):

\begin{enumerate}
  \item \textbf{Names mislead.} Account codes mislabel their contents
  (\texttt{atx}, \texttt{cler}, \texttt{craf}); dictionary descriptions
  rotate against their codes (\texttt{mach}/\texttt{nitc}/\texttt{trnp}).
  Any pipeline that classifies by prefix or label inherits these slips
  silently; ours refuses to, and both of the audit's largest apparent
  anomalies (R479~billion of phantom intermediates; a 125\,per\,cent
  value-added ``error'') dissolved into classification fixes.
  \item \textbf{Prose diverges from data.} A formula's printed signs do not
  reproduce the matrix they describe; the same cell is printed with two
  different values in two places. The data adjudicate.
  \item \textbf{Vintages drift.} The default download of a source (AFS 2019
  preliminary) is not the edition actually used (2019 revised); nothing in
  the record says so, and only a grid test can tell.
  \item \textbf{Inputs are unavailable.} One input is unpublished; its
  effect ($+16.1$\,per\,cent on import duties under the public proxy) is
  invisible until priced.
  \item \textbf{Adjustments go unrecorded.} A symmetric R596~million
  reconciling edit balances two accounts and appears nowhere in the record;
  inherited consumption shares import a different survey vintage's
  processing into a matrix whose documentation implies otherwise.
\end{enumerate}

To repeat the point made in Section~\ref{sec:results}: these are properties
of the \emph{medium}---prose describing spreadsheets---not of the authors,
whose record-keeping is far better than the norm. That is precisely what
makes the count suggestive: if the best-documented SAM in its class
carries five classes of silent divergence between record and artefact,
there is little reason to expect fewer in matrices published with no
dictionary, no formula note, and no distributed cell-level file---and for
those, the audit reported here could not even be attempted, because its
first step is a distributed matrix to compare against. One deep audit
cannot establish a distribution over the population of SAMs, and we do not
claim it does; what it establishes is that the divergences are not
hypothetical, and that detecting them required a machine-readable record
that most published SAMs do not supply.

The two further applications confirm that the taxonomy travels. Kenya
exhibits class~3 (the SAM's production side sits a uniform 1.0\,per\,cent
from the Economic Survey print, the signature of a provisional-accounts
vintage) and class~5 at national scale (the accounts' own
3.6-per-cent-of-GDP discrepancy is absorbed by the reconciliation with no
cell-level record of where). The Netherlands, with the cleanest data of
the three, still exhibits class~1: Eurostat's mixed-granularity category
lists and a NACE/ESA code collision would silently double-count under
prefix-based selection, and were caught only by the toolkit's tiling
assertion.

\subsection{From audit to construction workflow}
\label{sec:discussion:workflow}

The framework is not only a microscope; it is a factory. The same pipeline
that audits the benchmark \emph{builds} a SAM: scripts ingest and verify
the official tables, assemble the production and commodity blocks
mechanically, compute the institutional cells from published series,
apply survey-based share systems, and balance the result with a logged
adjustment. What remains for the economist is exactly the residue that
identities and obtainable data cannot determine---in this application, the
customs split, six sectors' capital shares, and the distributional detail
of household income. That inversion is the practical payoff: expert
attention stops being spread across ten thousand spreadsheet cells and
concentrates on the handful of judgments that genuinely require it, each
of which the framework forces into the open as a numbered rule. Updating
the matrix for a new benchmark year, or transferring it to a country with
comparable sources, reuses everything except those judgments.

\subsection{Implications: a standards proposal}
\label{sec:discussion:standards}

Two practices would have made most of this paper's detective work
unnecessary, and both are cheap. We state them as a proposal for producers
of SAMs---statistical offices, research institutes, and international
programmes alike:

\begin{enumerate}
  \item \textbf{Publish cell-level provenance.} Each cell of a released SAM
  should carry its source series, transformation, and any manual
  adjustment. The R596~million edit and the inherited consumption shares
  would then be declarations rather than discoveries, and reconciling
  interventions---a normal part of compilation---would be visible instead
  of latent.
  \item \textbf{Publish machine-readable construction rules.} The
  benchmark's technical note was nearly a program; had it been an actual
  one (as our transcribed formula table now is), the sign discrepancy could
  not have survived contact with the data, and replication would be a
  command rather than a project. Concordances, account dictionaries, and
  balancing targets belong in the same release.
\end{enumerate}

Neither requires changing a single methodological choice, and both are
by-products of building the matrix in code to begin with. Our toolkit is
offered as working infrastructure for exactly that: an open reference
implementation of registered sources, executable identities, explicit
classification, formula tables, diagnostic balancing, and workbook
generation, intended to be reused wholesale for other countries and years.

\subsection{Cost: a three-country ledger}

We report the development ledger factually and claim nothing causal about
labour saving from it. The South African case required registering
sources from four providers, one registered microdata access, and bulk
downloads worked around a non-functioning online query system; the Kenyan
case required locating condensed print tables inside a 322-page yearbook;
the Dutch case required three API calls and roughly 300 new lines of code
(70 of them a reader that now serves every ESA-transmitting country). The
claim we do make is conditional and testable: the marginal cost of the
\emph{next} SAM built on the toolkit is largely the cost of its
country-specific concordances and share systems, and the Dutch
application is a first operational indication of that claim at the macro
level. Across the three cases, publication format emerges as a major
determinant of marginal implementation effort---which is what the
standards proposal above is for.

\subsection{Limitations}

The deep audit inherits its benchmark's scope---one country, one year, one
SAM family---and the two further applications, while they demonstrate
transfer, are deliberately shallower: Kenya stops at structure and macro
controls because the machine-readable detail is not public, and the Dutch
exercise generates and balances a macro SAM, not a full multi-sector SAM with
distributional detail (that extension is mechanical for the production
core, which the API supplies at 65 industries, but household disaggregation
would require the national budget-survey work the South African case
illustrates). Six activities' capital splits and the item-level detail of the
household systems remain beyond the inputs we hold, and our
balanced household block is a disciplined proxy, not a reconstruction---
closing that gap requires the survey's item-level classifications, which
are obtainable and queued. The recovered choices (editions, measures,
signs) are unique within their tested grids, not proofs of intent; where
no grid candidate fits exactly---as in the capital split---the true basis
may lie outside the set we enumerated.

Two limits are structural rather than circumstantial, and neither is
removed by better data. The first concerns what the method is able to
detect. Every finding reported here was identified because it left an
arithmetic trace: an identity that did not close, a grid search returning
a single survivor, or a pair of gaps whose symmetry was unlikely to be
coincidental. Divergences that leave no such trace pass through this
framework without being detected. These include compensating errors,
adjustments that happen to land on a plausible value, and errors shared
between the artefact and the sources against which it is checked. The
audit therefore establishes that a matrix is reproducible from its stated
record, which is a narrower property than correctness. The comparison in
Section~\ref{sec:results:jrc} narrows this gap without closing it: an
independently constructed SAM for the same country-year agrees with ours
on every macro aggregate to within 0.8~per~cent, which is meaningful
corroboration, but it remains agreement between two constructions resting
on a shared official source rather than a second measurement of the same
economy. An error contained in the source record itself would not be
detected by any check reported here.

The second limit is that the framework records without evaluating. It
reports every adjustment factor, coverage share, and residual, but it
does not encode a threshold above which a matrix should be rejected, and
we do not propose one. The tolerable magnitude of a balancing adjustment
depends on the account, the country, and the intended use of the SAM, so
a threshold calibrated to three cases would carry more apparent authority
than the evidence supports. What the framework changes is the
availability of the question rather than its answer: because the factors
and residuals are published, a reader can apply whichever threshold their
application requires. Rejection criteria for the field would need to be
agreed among producers rather than set by a single implementation.

Finally, our own pipeline's mistakes---one of which we document---were
caught by the same assertions we advocate; we take that as the strongest
available evidence for building them in.
