\section{Background: the practice and its reproducibility problem}
\label{sec:related}

\subsection{Related work}

The SAM as an integrated accounting framework descends from
\citet{stone1962} and was consolidated as a planning tool by
\citet{pyattthorbecke1976} and \citet{pyattround1985}. The mechanics of
reconciling inconsistent accounts are equally mature: biproportional (RAS)
adjustment and its generalisations---GRAS for matrices carrying negative
entries \citep{junius2003,temurshoev2013}, KRAS where constraints
conflict or differ in reliability \citep{lenzen2009}---and the
cross-entropy estimation of \citet{robinson2001}, which remains the
standard reference for SAM balancing under scattered and unequally reliable
information. Compilation \emph{practice}, as distinct from compilation
mathematics, has its own literature: \citet{round2003} sets out the
lessons and unresolved difficulties of building SAMs for policy analysis,
and estimation manuals aimed at producers---most directly
\citet{mainar2018}, the manual behind the European Commission's own
African SAM programme---document the sequence of steps a compiler
follows. What these do not do, and what this paper is about, is make the
steps executable and the resulting artefact checkable against them.
Compilation practice for the supply--use core is codified in
the SNA and its handbooks \citep{un2009sna,un2018handbook,eurostat2008};
the analytics that consume SAMs are surveyed by \citet{millerblair2009}
and, for CGE databases, \citet{lofgren2002}. Operationally, most African
country SAMs are produced within a small number of institutional
pipelines---notably IFPRI's country programmes \citep{thurlow2021},
UNU-WIDER's South Africa series, of which the benchmark used here
\citep{vanseventer2023} is a leading example, and the Joint Research
Centre's Africa programme, which has published a 2019 South African SAM
disaggregated by province, gender, and formal--informal status
\citep{elmeligi2025}. South Africa's 2019 accounts have thus been
rendered as a SAM at least three times, independently and for different
analytical purposes. The cell-level audit reported here is against the
UNU-WIDER benchmark alone, but
Section~\ref{sec:results:jrc} compares our result with the Commission's
at the macro level---a reconciliation the field performs more rarely than
the existence of three renderings would suggest.

The paper's infrastructure claims sit at the junction of several
adjacent literatures, each of which has solved a piece of the problem.
Large international input--output databases are compiled under documented,
code-driven procedures---WIOD \citep{timmer2015}, EXIOBASE
\citep{stadler2018}, and Eurostat's FIGARO tables \citep{remond2019}---and
open-source tooling such as \texttt{pymrio} \citep{stadler2021} has made
their \emph{consumption} programmatic; but these systems document
methodology at the level of reports, not cell-level provenance, and none
is designed to let an outsider audit a single published matrix against
its own record. Machine-readable statistical standards (SDMX for
dissemination \citep{sdmx2021}; the W3C PROV model for data lineage
\citep{w3cprov2013}) supply exactly the vocabulary our standards proposal
needs, but have not been applied to SAM construction. And the norms of
reproducible computational research---\citet{peng2011},
\citet{sandve2013}, and for economics practice
\citet{gentzkow2014}---prescribe the workflow discipline the framework
implements, without the accounting-identity instrument that makes SAM
auditing unusually sharp. Within economics the paper joins the
transparency and replication agenda \citep{christensen2018}, with that
accounting twist: a SAM's identities give the auditor a powerful
instrument, because any undocumented intervention must either break an
identity or leave a biproportional fingerprint.

The closest existing systems are the input--output virtual laboratories,
of which the Australian Industrial Ecology Lab is the developed case
\citep{lenzen2014}: registered source ``data packs'', scripted assembly,
and constrained balancing under KRAS, built so that many analysts can
compile tables collaboratively and regenerate them. The virtual-laboratory
programme establishes most of what we assume here, and we claim no
priority over it. Two things differ, and they are differences of purpose
rather than sophistication. A virtual laboratory is built to
\emph{construct} tables, so its provenance record is oriented to the
build; our concern is the inverse operation---taking a matrix someone
else published, with its own prose record, and testing whether that
record reproduces it. And the laboratories serve environmental-economic
MRIO accounts, where the reference artefact is the database itself,
rather than the nationally published SAM whose documentation is the
object of audit. What we have not found is a published, cell-level
replication of a national SAM \emph{from its own stated sources}; if such
an audit exists we would welcome the correction, since the paper's
argument is strengthened, not weakened, by evidence that the exercise is
feasible more often than the record suggests.

\subsection{Why current workflows resist reproduction}
\label{sec:related:workflow}

The prevailing construction workflow---official tables copied into
spreadsheets, reconciled by hand, described afterwards in prose---fails
five requirements at once, none of them because practitioners are careless
(Table~\ref{tab:workflow}). \emph{Transparency}: judgments act on
individual cells but are reported, if at all, at the level of methods
paragraphs. \emph{Versioning}: a spreadsheet's history is its author's
memory; source vintages and revisions are rarely pinned.
\emph{Provenance}: a finished cell does not say which source, which
transformation, and which manual edit produced it. \emph{Auditability}: a
reader can check that the matrix balances, but not why any cell holds its
value. \emph{Reproducibility}: rebuilding the matrix from its stated
sources is a research project, not a command---as this paper demonstrates
by undertaking exactly that project. To our knowledge it is the first
full cell-level replication audit of a published national SAM from primary
sources; that such an audit is novel is itself evidence of the problem.

\begin{table}[ht]
\centering\small
\caption{Prevailing SAM construction workflow versus the framework used here.}
\label{tab:workflow}
\begin{tabularx}{\textwidth}{lXX}
\toprule
Requirement & Prevailing practice & This framework \\
\midrule
Transparency & Judgments described in prose, if at all & Classifications, formulas, and rules are machine-readable files \\
Versioning & Source vintages implicit; spreadsheet history unrecorded & Every input registered with vintage, access date, SHA-256 hash \\
Provenance & Finished cells carry no origin & Cells trace to source series, transformation, and logged adjustment \\
Auditability & Balance checkable; cell values not & Identities enforced as executable checks; adjustment log published \\
Reproducibility & Rebuild is a research project & Rebuild is a command (scripts 01--12 of the replication package) \\
\bottomrule
\end{tabularx}
\end{table}
